A quantum parallel Markov chain Monte Carlo

### 2021-11-29 main.tex
intro polished

%\emph{In the presence of these trade-offs, we seek to enjoy the benefits of parallel MCMC while (\textcolor{red}{a}) limiting the total amount of computing necessary and (\textcolor{red}{b})  retaining conventional parallel computing resources for secondary computational bottlenecks such as proposal generations and proposal density evaluations. Here, we assert that established quantum algorithms can help achieve these twin goals.}

Quantum computers use quantum mechanics to store information and manipulate data. By leveraging the idiosyncracies of quantum physics, these computers are able to deliver speedups over conventional computing for a relatively small number of computational problems. Some of these speedups are very large. Quantum computers can achieve logarithmic complexity for some classically linear time algorithms, turning $\order{N}$ into $\order{\log N}$.  Shor's quantum algorithm for integer factorization \citep{shor} and the HHL algorithm for linear system solving \citet{harrow2009quantum} deliver just such an \emph{exponential} speedup \citep{shor}. Other algorithms deliver a still impressive \emph{polynomial} speedup.  For example, the algorithms considered in the following (Section \ref{sec:qmin}) achieve quadratic speedups over conventional computing, turning  $\order{N}$ runtimes into $\mathcal{O}(\sqrt{ N})$.  
 Both the magnitude of quantum computing's successes and the relative rarity of those successes mean that there is a general interest in leveraging quantum algorithms for previously unconsidered computational problems.  We will show that---with the help of the Gumbel-max trick (Section \ref{sec:gm})---established quantum optimization techniques are actually useful for sampling from computationally expensive discrete distributions and apply this insight to achieving quadratic speedups for parallel MCMC.

In the following section, we provide a \emph{very} short introduction to a few of the basic elements of quantum computing. The interested reader may look to \citet{nielsen2002quantum} for a canonical introduction to quantum algorithms and their implementation.

>>>>

Quantum computers use quantum mechanics to store information and manipulate data. By leveraging the idiosyncracies of quantum physics, these computers are able to deliver speedups over conventional computing for a relatively small number of computational problems. Some of these speedups are very large. Quantum computers can achieve \emph{exponential} speedups over conventional computers for some computational tasks..  Shor's quantum algorithm for factoring an integer $N$ \citep{shor} is polynomial in $\log N$ compared to a super-polynomial classical optimum. The HHL algorithm for solving sparse $N$-dimensional linear systems \citep{harrow2009quantum} is $\order{\log N}$ compared to a classical $\order{N}$. Other algorithms deliver a still impressive \emph{polynomial} speedup.  For example, the algorithms considered in the following (Section \ref{sec:qmin}) achieve quadratic speedups over conventional computing, turning  $\order{N}$ runtimes into $\mathcal{O}(\sqrt{ N})$.  
 Both the magnitude of quantum computing's successes and the relative rarity of those successes mean that there is a general interest in leveraging quantum algorithms for previously unconsidered computational problems.  We will show that---with the help of the Gumbel-max trick (Section \ref{sec:gm})---established quantum optimization techniques are actually useful for sampling from computationally expensive discrete distributions and apply this insight to achieving quadratic speedups for parallel MCMC.

In the following section, we provide a \emph{very} short introduction to the most basic elements of quantum computing. The interested reader may look to \citet{nielsen2002quantum} for a canonical introduction to quantum algorithms and their implementation.
